Foreground Signatures and Mitigation in Neutral Hydrogen Power Spectra From The Reionization Epoch

### 2014-10-23 foreground_delay_spectrum_modeling.tex
from 6865031: ============
(*) Removed the histogram figure from git tracking
(*) Removed the corresponding commented portions relating to the histogram
figure
Minor edits to address collab comments. Mostly section 3

The MWA construction was completed in 2012 and, after commissioning, began its EoR observing program in 2013. In its final configuration the MWA is a 128--tile interferometer array capable of observing a 30.72~MHz instantaneous band anywhere in the range 80--300~MHz. Each tile is a phased array of 16 dipoles, each in the shape of a bow--tie. This yields a primary field of view $\gtrsim$~20\arcdeg~ wide and multiple secondary lobes. The array is arranged as a centrally condensed core of $\sim$~300~m --- there are many spacings in the range 5--50~m --- and a radial density that falls off as the inverse of the radius, with the longest baselines at 3~km \citep{bea12}. Here we focus on antenna spacings $|\vec{b}| \le 200$~m (spatial scales relevant to reionization) and a frequency range 170--200~MHz. The center frequency of 185~MHz corresponds to a redshift, $z\simeq 6.68$.

@@

The data are flagged\footnote{\url{http://sourceforge.net/p/aoflagger}} for interference \citep{off12,off10}, removing 3\% of the data and averaged in time  and frequency from the raw 0.5~s, 40~kHz to 2s, 80~kHz. These data are then calibrated to a simulation of the sky containing 2420 point--like objects selected from the MWA Commissioning Survey (MWACS; Hurley-Walker et al. 2014, submitted). % MWACS was compiled using observations taken with subsets of the array during the final buildout and commissioning;
It has a flux density limit of 25~mJy and a declination range of -12\arcdeg~ to -40\arcdeg~ evenly covering the field of view of the observations reported here. The objects used in calibration are selected to lie inside the 5\% contour of the primary lobe of the tile power--pattern. The calibration algorithm -- based on forward modeling software by \citet{sul12} -- computes complex gain solutions per channel per antenna averaged to two minute intervals. The solutions are fairly low signal--to--noise so we iteratively project along the antenna and frequency dimensions to capture the relatively independent passband and antenna--to--antenna variation. First, we average the channel gains over all antennas to obtain a high signal--to--noise measurement of the bandpass. After applying this single passband, we do a second round of calibration and fit second and first order polynomials for amplitude and phase respectively for each antenna. This flattens any residual variation in bandpass and removes small phase slopes due to variations in cable delay. Finally, we fit for an additional phase known to be caused by reflections in a few 150~m length cables. %possibly could add elaboration, if I knew exactly which version of the cal was done here. doesn't really matter

@@

We obtain the delay spectrum of these calibrated visibilities taking the delay transform of each baseline's spectrum (equation~\ref{eqn:delay-transform}) choosing $W'_f(f)$ to be a {\it Blackman--Harris} window function. The spectrum is multiplied by the additional weight of the flagged channels which occur rarely for interference and regularly every 1.28~MHz where the edges of coarse channels are known to have a small aliasing effect. In delay--space, these weights translate into a convolution by a grated point spread function (PSF) with gratings at multiples of $\tau=0.78\,\mu$s. We deconvolve this PSF using a one dimensional CLEAN algorithm \citep{tay99} as described for the delay axis by \citet{par09,par12}. The CLEAN procedure iteratively finds and subtracts peak values convolved by the Fourier transform of the weights. We limit the selection of peaks to modes inside the horizon delay limit, corresponding to smooth spectrum objects in the visible sky hemisphere.

@@

% Figure~\ref{fig:data-sim-ratio} shows normalized histograms of the distribution of $\log_{10}\rho$ for the {\it off--zenith} (top panel) and {\it zenith} (bottom panel) pointings respectively inside the {\it foreground wedge}. The median absolute deviation from these distributions corresponds to $\sim$~90\% fractional difference between data and modeling on average in either case. This is in line with expectations when the aforementioned uncertainty in foreground models, thermal noise fluctuations in measurements, and uncertainties in antenna power pattern due to random delay fluctuations are taken into account. Currently, we are significantly limited by the unavailability of all--sky foreground models at the observing frequency of 185~MHz.

% \begin{figure}[htb]
% \centering
% \includegraphics[width=\linewidth]{figures/v1_0/delta_array_histogram_wedge_sim_data_log_ratio_0.3m_ground_custom_gaussian_FG_model_asm_all_sky_nside_64_Tsys_95.0K_185.0_MHz_30.7_MHz_bhw2.0.eps}
% \caption{Normalized histograms of logarithm of ratios of observed to simulated visibilities restricted to the {\it foreground wedge} -- i.e., within the white lines of Figure \ref{fig:sim_data} -- for the {\it off--zenith} (top) and {\it zenith} (bottom) pointings. Median absolute deviation of $\log_{10}\rho$ is $\sim 0.28$ in both cases indicating a fractional difference of $\sim 90$\%. A value of zero indicates perfect agreement ($\rho=1$). \label{fig:data-sim-ratio}}
% \end{figure}

>>>>

MWA construction was completed in 2012 and, after commissioning, began its EoR observing program in 2013. In its final configuration the MWA is a 128--tile interferometer array capable of observing a 30.72~MHz instantaneous band anywhere in the range 80--300~MHz. Each tile is a phased array of 16 dipoles, each in the shape of a bow--tie. This yields a primary field of view $\gtrsim$~20\arcdeg~ wide and multiple secondary lobes. The array is arranged as a centrally condensed core of $\sim$~300~m --- there are many spacings in the range 5--50~m --- and a radial density that falls off as the inverse of the radius, with the longest baselines at 3~km \citep{bea12}. Here we focus on antenna spacings $|\vec{b}| \le 200$~m (spatial scales relevant to reionization) and a frequency range 170--200~MHz. The center frequency of 185~MHz corresponds to a redshift, $z\simeq 6.68$.

@@

The data are flagged\footnote{\url{http://sourceforge.net/p/aoflagger}} for interference \citep{off12,off10}, removing 3\% of the data and averaged in time  and frequency from the raw 0.5~s, 40~kHz to 2s, 80~kHz. These data are then calibrated to a simulation of the sky containing 2420 point--like objects selected from the MWA Commissioning Survey (MWACS; Hurley-Walker et al. 2014, submitted). % MWACS was compiled using observations taken with subsets of the array during the final buildout and commissioning;
It has a flux density limit of 25~mJy and a declination range of -12\arcdeg~ to -40\arcdeg~ evenly covering the field of view of the observations reported here. The objects used in calibration are selected to lie inside the 5\% contour of the primary lobe of the tile power--pattern. The calibration algorithm -- based on forward modeling software by \citet{sul12} and the calibration method described by \citet{salvini2014fast} -- computes complex gain solutions per channel per antenna averaged to two minute intervals. The solutions are fairly low signal--to--noise so we iteratively average along the antenna and frequency dimensions to capture the relatively independent passband and antenna--to--antenna variation. First, we average the channel gains over all antennas to obtain a high signal--to--noise measurement of the bandpass. After applying this single passband, we do a second round of calibration and fit second and first order polynomials for amplitude and phase respectively for each antenna. This flattens any residual variation in bandpass and removes small phase slopes due to variations in cable delay. Finally, we fit for an additional phase known to be caused by small reflections in subset of cables. %possibly could add elaboration, if I knew exactly which version of the cal was done here. doesn't really matter

%@article{salvini2014fast,
%  title={Fast gain calibration in radio astronomy using alternating direction implicit methods: Analysis and applications},
%  author={Salvini, Stefano and Wijnholds, Stefan J},
%  journal={arXiv preprint arXiv:1410.2101},
%  year={2014}
%}

@@

We obtain the delay spectrum of these calibrated visibilities by taking the delay transform of each baseline's spectrum (equation~\ref{eqn:delay-transform}) choosing $W'_f(f)$ to be a {\it Blackman--Harris} window function. The spectrum is multiplied by the additional weight of the flagged channels which occur rarely for interference and regularly every 1.28~MHz where the edges of coarse channels are known to have a small aliasing effect. In delay--space, these weights translate into a convolution by a grated point spread function (PSF) with gratings at multiples of $\tau=0.78\,\mu$s. We deconvolve this PSF using a one dimensional CLEAN algorithm \citep{tay99} as described for the delay axis by \citet{par09,par12}. The CLEAN procedure iteratively finds and subtracts peak values convolved by the Fourier transform of the weights. We limit the selection of peaks to modes inside the horizon delay limit, corresponding to smooth spectrum objects in the visible sky hemisphere.

@@
